---
title: "mbts"
author: "SW"
date: "1/31/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source('gen_truth_class_pois.R')
```

## Generating Simulation Data

To generate simulated data:

```{r}
set.seed(1)

tbl <- gen_table()
```

The parameters in this function and their default values are:

* fl_sig=0: floor of the arima signal normalization
* w_sig=6: roof of the arima signal normalization
* fl_bg=-6: floor of the background normalization
* w_bg=6: roof of the background normalization
* bg_disp_mu=0: background noise poisson distribution mean
* bg_disp_sigma=1: background noise poisson distribution sd
* sig_disp_mu1=0: arima shared noise poisson distribution mean
* sig_disp_sigma1=0: arima shared noise poisson distribution sd
* sig_disp_mu2=0: signal taxa specific noise poission distribution mean
* sig_disp_sigma2=1: signal taxa specific noise poission distribution sd
* n_sig=10: number of arima signals
* n_clust=10: number of taxa in an arima signal
* n_tax_sig=1: number of taxa with FINAL signal (in beta atm)
* n_bg=700: number of background taxa (this + signal columns in output data)
* len_arima=1000: length of arima signal, needs to be larger than window
* len_ts=500: length of the time series (row size of output table)
* len_signal=300) # length of the shared signal

The output of gen_table() is the final simulated abundance table:

```{r}
tbl[1:5,1:5]
```

The column names reflect whether the features are signal or background (the noise):

```{r}
head(colnames(tbl))
tail(colnames(tbl))
```

The table belongs to the class mbts:

```{r}
class(tbl)
```

and it has a set of methods:

```{r}
methods(class='mbts')
```

There are two plotting methods. plot_sig() plots **examples** of the arima signal. This isn't the same signal used to generate the simulated data; instead, it's a set of similar signals obtained using the same arima parameters.

```{r}
plot_sig(tbl,n=4)
```

plot_sim plots the simulated data, where the blue shading shows the aligned, shared signal across all taxa (clusters):

```{r}
plot_sim(tbl)
```

The class also has a set of attributes with some of the intermediate data:

```{r}
names(attributes(tbl))
```

signals stores the underlying, untouched arima signals. For this example, there were 10 taxa (clusters), so this the signals slot will be a list of length 10:

```{r}
signals <- attr(tbl,'signals')
length(signals)
```

If we look at the first taxon, we can see that each slot in this list contains:

* the time series signal before noise (the pure signal)
* the time series signal after noise
* the final time series after timeshifting and adding the secound round of noise
* the indexes used to timeshift the signal

```{r}
taxa_1 <- signals[[1]]
names(taxa_1)

head(taxa_1$timeseries_pure)
head(taxa_1$timeseries)
taxa_1$cluster[1:5,1:5]
```

And we can plot these:

```{r}
qplot(seq_along(taxa_1$timeseries_pure),
      taxa_1$timeseries_pure,
      geom='line',xlab='',ylab='')

qplot(seq_along(taxa_1$timeseries),
      taxa_1$timeseries,
      geom='line',xlab='',ylab='')

data.frame(time=1:nrow(taxa_1$cluster),taxa_1$cluster) %>%
  gather(taxon,count,-time) %>%
  mutate(taxon=as.integer(gsub('X','',taxon))) %>%
  ggplot(aes(time,count)) +
  geom_point(size=.5) +
  geom_line(alpha=.5) +
  facet_grid(taxon~.) +
  theme_classic()
```

If we needed to access the background distribution, we can do the following:

```{r}
background <- attr(tbl,'background')
background[1:5,1:5]
```

And if we needed to access the parameters we originally passed to generate the simulation:

```{r}
attr(tbl,'params')
```
